{\large Gravitational wave sources in our Galactic backyard:}\\{\normalsize Models for detecting BHBH, BHNS and NSNS binaries with LISA

### 2021-11-26 paper/tex/sections/results.tex
results I proofread

For reference, we show dotted lines to indicate where a hypothetical binary system would reside assuming a given distance from earth (diagonal lines) and a fixed remaining inspiral time (vertical lines). For each line we assume the binary is circular and has a chirp mass equal to the average of the sample ($\avg{\mathcal{M}_c}$, annotated in each panel), unless indicated otherwise. We also overplot the LISA verification binaries (star symbols, \citealt{Kupfer+2018}).

@@

In the lower three panels, we show the contributions of the different types of sources. Comparing them, one can observe the shift in the frequency at which the peak is located, at $0.3 \unit{mHz}$, $0.7 \unit{mHz}$ and $1 \unit{mHz}$ for BHBH, BHNS and NSNS systems respectively. This is a result of the difference in chirp mass. A higher mass system can emit at lower frequency and still produce the minimum SNR needed for detection. We note that this effect can be used to distinguish the heavier DCOs that we discuss in this work from WDWD systems, at least probabilistically \citep[e.g.][]{Sesana+2020}. In the same way, this also explains the offset in frequency between the population of sources we predict and the LISA verification binaries.

Inspecting the dotted reference lines, we note that the peak of the density distribution of observable sources coincides with the location expected for circular systems at $8 \unit{kpc}$, which is the distance to the centre of the Milky Way. As can be seen best in the lower panels, the reference lines for $0.1$ and $30 \unit{kpc}$ enclose the majority of systems as expected given the dimensions of our Milky Way.

There is a distinct subpopulation of binaries, most clearly visible in the lower panels as an offshoot that extends downwards to ASDs of $10^{-19} \unit{Hz}^{-1/2}$, especially around $2 \unit{mHz}$. This offshoot is almost uniquely composed of eccentric binaries, as can be seen in Fig.~\ref{fig:dcos_on_sc_ecc_col} in our appendix, which shows a similar figure but colouring individual systems by their eccentricity. This can also be seen, albeit indirectly, from the contours shown in the bottom panels of Fig.~\ref{fig:dcos_on_sc}, which encompass 90\% of the \textit{circular} sources in each population. This contour does not include the offshoot. We conclude that eccentric sources occupy a very different region in this diagram.

@@

\paragraph{Orbital Frequency} The orbital frequency distributions for BHBHs, BHNSs and NSNSs (Fig.~\ref{fig:fiducial_pdf_distributions}a) peak at progressively increasing frequencies. As already mentioned above, this is because a higher mass DCO at the same distance and eccentricity requires a lower frequency to produce the same signal-to-noise ratio and thus be detected. The distributions appear nearly symmetric, but closer inspections shows that the left hand side is more populated, which can be seen most clearly in the curve for the BHBHs. This is due to the contribution of highly eccentric binaries, which are most abundant in the BHBH population. These systems are still detectable by LISA, despite their low orbital frequency, as the high eccentricity means that the majority of the GW signal is emitted at higher harmonics, where LISA is more sensitive.

@@

We also note that our BH mass distribution extends down below $5\unit{M_{\odot}}$ to $2.5 \unit{M_{\odot}}$ which is our fiducial maximum neutron star mass. The BHs in this simulation fill the so-called ``lower mass gap'' marked as a grey band \citep{Ozel+2010,Farr+2011}, see also \citet{Shao+2021} who recently pointed this out. This prediction is sensitive to adopted model for fallback during the SN explosion as we discuss this further in Section~\ref{sec:lower_mass_gap}

@@

We find that detectable BHNSs have even more unequal mass ratios. Moreover, the mass ratio distribution is bimodal, where the two peaks arise from two distinct formation scenarios. Around two thirds of detectable BHNSs experience at least one common-envelope event, whilst the last third are formed through only stable mass transfer. The first peak at $q = 0.18$ is from systems that experience at least one common-envelope phase and occurs at the expected mass ratio, which approximately follows the mean BH mass ($\sim 6.5 \unit{M_\odot}$) and NS mass ($\sim 1.2 \unit{M_\odot}$). Yet we also see a second peak at higher mass ratios around $q = 0.34$, which arises from the fraction of the population that underwent only stable mass transfer phases. The stability of mass transfer depends on the mass ratio, preferentially forming systems with more equal masses, i.e.\ at higher $q$.

@@

For systems with eccentricities higher than $e\gtrsim0.3$, most gravitational wave energy is emitted in higher harmonics. Such systems are more rare, but we find them to be significant among the BHBH population, where they account for \BHBHHighlyEccentric{} of systems. Detectable BHBHs in our simulation (and, in particular, those that are eccentric) are primarily systems formed through the stable mass transfer channel (see Section~\ref{fig:formation_channels}). These systems are still relatively wide (compared to those formed through the CE channel) immediately prior to formation of the second BH, which makes them more easily affected by kicks. If the kick is oriented roughly in opposite direction to the orbital motion and has a velocity that is of similar magnitude as the orbital velocity, it will lead to the formation of a highly eccentric system.

@@

\subsection{Distribution in the Milky Way}\label{sec:mw_detectable_distribution}
For our Milky way model we consider three different components, a low-\achem~(``thin'') disc, an older high-\achem~(``thick'') disc and a bulge/bar. In Table~\ref{tab:component_fractions} we summarise the number of detections originating from each of these components. Despite the fact that only $42.5\%$ of systems are formed in the low-\achem~disc, we find that $86\%$ of the detections originate from this component. This is because most detectable systems were formed relatively recently (see Fig.~\ref{fig:fiducial_pdf_distributions}e) and so the high-\achem~disc and bulge are effectively too old to contribute many detectable systems. Nevertheless, we do find a significant fraction of detectable systems originate in the high-\achem~disc and bulge, indicating that it is still important to include these components, as was ignored in some earlier works (see Sec.~\ref{sec:compare_studies}).

@@

The distribution for NSNS stands out by being flatter, making them more common nearby and, surprisingly, also at larger distances. This may seem counter intuitive as one might naively expect the less massive NSNS systems would not be observable out to larger distances than the more massive BHNS and NSNS systems. To understand the differences we need to consider not only the mass distribution of binaries, but also their eccentricity and frequency distributions. Since, each parameter contributes to the calculation of the SNR (and thus affects the maximum distance at which systems can be detected).

@@

The fraction of detectable DCOs that formed through different formation channels in other model variations are shown Fig.~\ref{fig:formation_channels}, where the first column corresponds to the fiducial model described above.

>>>>

For reference, we show dotted lines to indicate where a hypothetical binary system would reside assuming a given distance from earth (diagonal lines) and a fixed remaining inspiral time (vertical lines). For each line we assume the binary is circular and has a chirp mass equal to the average of the sample ($\avg{\mathcal{M}_c}$, annotated in each panel). We also overplot the LISA verification binaries (star symbols, \citealt{Kupfer+2018}).

@@

In the lower three panels, we show the contributions of the different types of sources. Comparing them, one can observe the shift in the frequency at which the peak is located, at $0.3 \unit{mHz}$, $0.7 \unit{mHz}$ and $1 \unit{mHz}$ for BHBH, BHNS and NSNS systems respectively. This is a result of the difference in chirp mass. A higher mass system can emit at lower frequency and still produce the minimum SNR needed for detection. We note that this effect can be used to distinguish the heavier DCOs that we discuss in this work from WDWD systems, at least probabilistically (see Sec.~\ref{sec:WDWD_distinguish}). In the same way, this also explains the offset in frequency between the population of sources we predict and the LISA verification binaries.

Inspecting the dotted reference lines, we note that the peak of the density distribution of observable sources coincides with the location expected for circular systems at $8 \unit{kpc}$, which is the distance to the centre of the Milky Way. As can be seen best in the lower panels, the reference lines for $0.1$ and $30 \unit{kpc}$ enclose the majority of systems, as expected given the dimensions of the Milky Way.

There is a distinct subpopulation of binaries, most clearly visible in the lower panels as an offshoot that extends downwards to ASDs of $10^{-19} \unit{Hz}^{-1/2}$, especially around $2 \unit{mHz}$. This offshoot is almost uniquely composed of eccentric binaries, as can be seen in Fig.~\ref{fig:dcos_on_sc_ecc_col}, which shows a similar figure but colouring individual systems by their eccentricity. This can also be seen, albeit indirectly, from the contour lines shown in the bottom panels of Fig.~\ref{fig:dcos_on_sc}, which encompass 90\% of the \textit{circular} sources in each population. This contour does not include the offshoot. We conclude that eccentric sources occupy a very different region in this diagram.

@@

\paragraph{Orbital Frequency} The orbital frequency distributions for BHBHs, BHNSs and NSNSs (Fig.~\ref{fig:fiducial_pdf_distributions}a) peak at progressively increasing frequencies. As mentioned in Sec.~\ref{sec:dcos_on_sc}, this is because a higher mass DCO at the same distance and eccentricity requires a lower frequency to produce the same signal-to-noise ratio and thus be detected. The distributions appear nearly symmetric, but closer inspections shows that the left hand side is more populated, which can be seen most clearly in the curve for the BHBHs. This is due to the contribution of highly eccentric binaries, which are most abundant in the BHBH population. These systems are still detectable by LISA, despite their low orbital frequency, as the high eccentricity means that the majority of the GW signal is emitted at higher harmonics, where LISA is more sensitive.

@@

We also note that our BH mass distribution extends down below $5\unit{M_{\odot}}$ to $2.5 \unit{M_{\odot}}$ which is our fiducial maximum neutron star mass. The BHs in this simulation fill the so-called ``lower mass gap'' marked as a grey band \citep{Ozel+2010,Farr+2011}, see also \citet{Shao+2021} who recently also pointed this out. This prediction is sensitive to adopted model for fallback during the SN explosion as we discuss this further in Section~\ref{sec:lower_mass_gap}.

@@

We find that detectable BHNSs have even more unequal mass ratios. Moreover, the mass ratio distribution is bimodal, where the two peaks arise from two distinct formation scenarios. Around two thirds of detectable BHNSs experience at least one common-envelope event, whilst the last third are formed through only stable mass transfer. The first peak at $q = 0.18$ is from systems that experience at least one common-envelope phase and occurs at the expected mass ratio, which approximately follows the mean BH mass (${\sim}6.5 \unit{M_\odot}$) and NS mass (${\sim}1.2 \unit{M_\odot}$). Yet we also see a second peak at higher mass ratios around $q = 0.34$, which arises from the fraction of the population that underwent only stable mass transfer phases. The stability of mass transfer depends on the mass ratio, preferentially forming systems with more equal masses, i.e.\ at higher $q$.

@@

For systems with eccentricities higher than $e\gtrsim0.3$, most gravitational wave energy is emitted in higher harmonics. Such systems are more rare, but we find them to be significant among the BHBH population, where they account for \BHBHHighlyEccentric{} of systems. Detectable BHBHs in our simulation (and, in particular, those that are eccentric) are primarily systems formed through the stable mass transfer channel (see Fig.~\ref{fig:formation_channels}). These systems are still relatively wide (compared to those formed through the CE channel) immediately prior to formation of the second BH, which makes them more easily affected by kicks. If the kick is oriented roughly in opposite direction to the orbital motion and has a velocity that is of similar magnitude as the orbital velocity, it will lead to the formation of a highly eccentric system.

@@

\subsection{Distribution in the Milky Way}\label{sec:mw_detectable_distribution}
For our Milky way model we consider three different components, a low-\achem~(``thin'') disc, an older high-\achem~(``thick'') disc and a bulge/bar (see Sec.~\ref{sec:galaxy_synthesis}). In Table~\ref{tab:component_fractions} we summarise the number of detections originating from each of these components. Despite the fact that only $42.5\%$ of systems are formed in the low-\achem~disc, we find that $86\%$ of the detections originate from this component. This is because most detectable systems were formed relatively recently (see Fig.~\ref{fig:fiducial_pdf_distributions}e) and so the high-\achem~disc and bulge are effectively too old to contribute many detectable systems. Nevertheless, we do find a significant fraction of detectable systems originate in the high-\achem~disc and bulge, indicating that it is still important to include these components, as was ignored in some earlier works (see Sec.~\ref{sec:compare_studies}).

@@

The distribution for NSNS stands out by being flatter, making them more common nearby and, surprisingly, also at larger distances compared to the stellar density. This may seem counter intuitive as one might naively expect the less massive NSNS systems would not be observable out to larger distances than the more massive BHNS and NSNS systems. To understand the differences we need to consider not only the mass distribution of binaries, but also their eccentricity and frequency distributions. Since, each parameter contributes to the calculation of the SNR (and thus affects the maximum distance at which systems can be detected).

@@

The fraction of detectable DCOs that are formed through different formation channels in the other model variations are shown Fig.~\ref{fig:formation_channels}, where the first column in the plot corresponds to the fiducial model that we described above.
